Quantitative Benchmarking for Automatically Generated Conjectures

### 2018-03-16 TEBenchmarkPaper/article.tex
Background

Equation \ref{eq:mapreduce} justifies a rewrite rule for splitting a single
$\Map$ call into multiple independent calls dealing with different sections of a
list. Such rules can be useful optimisations, for example in a distributed
context where since each call can be evaluated in parallel (this is the basis of
the ``map/reduce'' programming paradigm).

All of these example use cases require the ability to discover theorems about
some particular theory. This is a hard problem in general, but presents
opportunities for automation due to the precise, symbolic nature of the domain.

@@

{\bf Utility} is the relevance or usefulness of a conjecture to the
user's particular task. For example, if we want to find optimising
rewrite rules such as equation 1, then utility would include whether
or not a conjecture justifies a rewrite rule, the difference in
resource usage of the expressions involved, and how common those
expressions are in real usage.

@@

The latter tools are described in more detail in section~\ref{section:mte},
but all use testing to check for counterexamples, which ensures results are
empirically plausible and applicable (conditions are satisfiable, types are
inhabited, etc.). They also use a small-to-large search order, ensuring more
easily-comprehensibile conjectures are explored first.

@@

We believe that ATF and MTE share many goals and would benefit from a
closer alignment: both fields are highly specialised and by joining
forces (while being explicit about differences) the techniques being
developed would have greater impact. This closer alignment is already
beginning to take place. The main commonalities between MTE and ATF
are that both are interested in the same aspects of mathematical
thinking, aiming to automatically construct and evaluate elements of a
mathematical theory, including concepts, conjectures, theorems, axioms
and examples. Both strands contrast themselves with Automated Theorem
Proving. The main historical differences between these two approaches
seem to be that in MTE:

\begin{itemize}
\item the main proponents define themselves as mathematicians (they
  hold a PhD and have experience in research mathematics);
\item the primary motivation is to support mathematicians;
\item systems tend to be user-interactive;
\item systems tend to be specific to mathematics,
\end{itemize}

while in ATF:

\begin{itemize}
\item the main proponents define themselves as AI researchers (they
  often have a mathematics background up to Masters level, but their
  PhD and research experience is in AI);
\item the primary motivation is to extend AI techniques; a secondary
  motivation is to produce interesting new mathematics;
\item systems tend to be fully automated;
\item systems can often be applied to non-mathematical domains.
\end{itemize}

>>>>

Equation~\ref{eq:mapreduce} justifies a rewrite rule for splitting a single
$\Map$ call into multiple independent calls dealing with different sections of a
list. Such rules can be useful optimisations, for example in a distributed
context where each call can be evaluated in parallel (this is the basis of the
``map/reduce'' programming paradigm).

All of these example use cases require the ability to discover theorems about
some particular definitions. This is a hard problem in general, but presents
opportunities for automation due to the precise, symbolic nature of the domain.

@@

{\bf Utility} is the relevance or usefulness of a conjecture to the
user's particular task. For example, if we want to find optimising
rewrite rules such as equation~\ref{eq:mapreduce}, then utility would include
whether or not a conjecture justifies a rewrite rule, the difference in
resource usage of the expressions involved, and how common those
expressions are in real usage.

@@

The latter tools are described in more detail in $\S$\ref{section:mte},
but all use testing to check for counterexamples, which ensures results are
empirically plausible and applicable (conditions are satisfiable, types are
inhabited, etc.). They also use a small-to-large search order, ensuring more
easily-comprehensibile conjectures are explored first.

@@

We believe that ATF and MTE share many goals and would benefit from a
closer alignment: both fields are highly specialised and by joining
forces (while being explicit about differences) the techniques being
developed would have greater impact. This closer alignment is already
beginning to take place. The main commonalities between MTE and ATF
are that both are interested in the same aspects of mathematical
thinking, aiming to automatically construct and evaluate elements of a
mathematical theory, including concepts, conjectures, theorems, axioms
and examples. Both strands contrast themselves with Automated Theorem
Proving. The main historical differences between these two approaches seem to be
that MTE was initiated by mathematicians as an attempt to support existing
mathematical research, perhaps with user interaction; while ATF is a product of
AI researchers trying to extend AI techniques, with full automation and often
applications to non-mathematical domains.
